---
title: "Group_07_Analysis"
author: "Group 7"
output:
  pdf_document:
    latex_engine: xelatex
    number_sections: yes
  html_document:
    df_print: paged
fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA)
```

```{r libraries}
library(tidyverse)
library(moderndive)
library(skimr)
library(kableExtra)
library(gridExtra)
library(broom)
library(janitor)
library(ggplot2)
library(GGally)
library(sjPlot)
```

```{r data}
film <- read.csv("dataset7.csv") 
# Split output Y into 2 groups >7 and <=7  
film <- film %>% 
  mutate(Rate = ifelse(film$rating >7 , ">7" ,"<=7"))%>%
  mutate_at(vars(Rate ) , funs(factor))
# Find NA observations 
# There are 92 missing data 

film %>% 
   select(rating) %>%
  summary()

ggplot(film, aes(x=rating)) + 
   geom_histogram(col="white",fill="wheat4")+
   ggtitle("Density Of Rating")+
   labs(y="frequency")+
   scale_x_continuous(name="Rating", limits=c(0, 10),n.breaks = 10)+
   theme_light()

```


```{r}
  
# Genre Variable #
#================#

# Contingency Table 


film %>% 
     tabyl(Rate, genre) %>% 
     adorn_percentages() %>% 
     adorn_pct_formatting() %>% 
     adorn_ns() 

# check varaibility 
ggplot(film , aes(x=genre)) + 
   geom_bar(col= "white",fill="SkyBlue3") + 
   geom_text(aes(label=..count..),stat="count", vjust=1.6, color="white", size=3.5)+
   theme_light()
# also given the table 
film %>% 
   select(genre) %>% 
   table()

     # Action   Animation      Comedy Documentary       Drama     Romance       Short 
     #    698         165         582         136         684          16         106 

ggplot ( film , aes(x= Rate, y= ..prop.., group= genre,fill=genre)) + 
   geom_bar(col ="white" , position="dodge",stat = "count") + 
   scale_fill_discrete()+
   labs(x="Rate" , "y=Proportion")+
   theme_light()
 
```
 
```{r}
  ggplot(film, aes(x = rating)) +
    geom_density(aes(color = genre)) +
    theme_light()
```
From this density gragh we can see that Short,  Documentary,Comedy and Animation films are more likely to get rates over 7.5 compared to rates 
lower than 7 ( 5 to be exact )

We notice that the majority of films are within one of the 3 most dominant categories 
Action , Comedy and Drama .
On the opposite Animation Documentary and Short films have relatively close frequencies . 
And Romance films are comparatively very rare among our list of films ( only 16 films where Romantic among all 2387)

```{r} 
# Romance films 
summary(film[film$genre=="Romance",]$rating)
```

Given the barchart we notice that in our dataset all Romance films have ratings below 7. 
This might indicate that Romance is not a strong explanation of rating. 
 
 
```{r}
# Budget # 
#=========#
ggplot( film , aes( x= Rate , y= budget,fill=Rate )) + 
   geom_boxplot() +
   scale_fill_manual(values=c("coral2", "lightblue"))+
   labs(x="Ratings" , y =paste("Budget (in $ 10 ^3's"))+
   theme_light()


```

```{r}
# Length # 
#==========# 

summary(film$length)  # There are 92 missing data 

# Comparing the 3rd Quartile and Maximum value ( 100, #399 respectively ) we see that length has many outliers

dim(na.omit(film[  film$length >100 & film$Rate==">7",] ))[1]

# Check density of length 
ggplot(film , aes(x=length)) + 
   geom_bar()+
   theme_light()

# Rate ~ Length 
ggplot(film , aes(x= Rate , y= scale(length), fill = Rate)) + 
   geom_boxplot() +
   scale_fill_manual(values=c("coral3", "lightblue"))+
   labs(x="Ratings" , y ="Length (in minutes)")+
   theme_light()


# Year # 
#=======# 
ggplot( film , aes(x = Rate , y = year )) + 
  geom_boxplot()
  



```

```{r}

# Positive Votes # 
#=================#

# Check variability of Votes variable 
ggplot(film , aes(x=scale(votes),y=rating))+ 
   geom_jitter(col= "aquamarine2",width=0.5)+
   labs(x="Votes ( scaled) " , y="Rating ( 1-10) " , title = "Rating of films based on Positive Votes ")+
      theme_light()

 ggplot(film , aes(y= log(votes) , x= Rate , fill=Rate )) + 
   geom_boxplot() + 
   scale_fill_discrete()+
   labs(x="Rate" , y = "Number of Positive Votes")+
   theme_light()
```

```{r}
    # Year # 
#=================#
ggplot(film , aes(x = (year) , group= Rate , fill=Rate)) + 
   scale_fill_manual(values = c("coral3" , "steelblue"))+
   geom_bar()


```

```{r}
# Correlation between the explanatory variables 
ggplot(No_Rom_films , aes(x=genre , y=length )) + 
   geom_boxplot(fill="HotPink3")+
      theme_light()

corr <- data.frame(lapply(film[,-c(1,6)], as.integer))
corr <- corr %>% 
   mutate_all(scale)
# Plot the graphg
library(GGally)
ggcorr(corr,
    method = c("pairwise", "spearman"),
    nbreaks = 6,
    hjust = 0.8,
    label = TRUE,
    label_size = 3,
    color = "grey50")
``` 
We see in general films' duration are roughly in same range except Animation and Short films. Both have comparatively smaller length. 

```{r model}

# Data frame with Romance genre removed 
No_Rom_films <- film[!(film$genre == "Romance"),]

# Remove missing values 
No_Rom_films <- na.omit(No_Rom_films)

# Now genre is categorical while length is numerical 
 model <- glm( Rate ~ genre + length , data = No_Rom_films , 
               family = binomial(link = "logit"))
 summary(model)

```


# Introduction {#sec:Intro}
IMDb is a platform that provides information , ratings and reviews on films, movies and many other streaming content. 
Our group was given a dataset taken from this databese, which has records of 2387 films released from  ` r min(film$year) ` to `r max(film$year)`
Each film has a unique identification number (id) and some other measurments related to it. Those are : 
\begin {itemize} 
\item $ Year_i $   : the year the film was released at cinemas 
\item $ Length_i $ : duration of film in minutes 
\item $ Budget_i$  : Budget for film production in$ \$ 10^ 5  $'s
\item $ Vote_i $   : Number of positive votes
\item $ Genre_i $  : Genre of film
\item $ Rating_i $ : IMDb rating from 1 to 10 

\end {itemiza}





# Exploratory Data Analysis {#sec:EDA}

```{r summary}
my_skim <- skim_with(base = sfl(n = length))
film%>%
  select(-c(1:2,6,8)) %>% 
  my_skim() %>% 
  transmute(Variable=skim_variable, Mean=numeric.mean, SD=numeric.sd,
            Min=numeric.p0, Q1=numeric.p25, Median=numeric.p50, Q3=numeric.p75,Max=numeric.p100,
            IQR = numeric.p75-numeric.p50) %>% 
  kable(caption = '\\label{tab:summary}Summary statistics of variables in the data set.' ,
        booktabs = TRUE, linesep = "", digits = 3) %>%
  kable_styling(font_size = 10, latex_options = "hold_position")
```
Taking a glance at the summaries of the variables we see notice that length variable has many outliers , and that is given by 
the big difference between its maximume value and its third quartile value ,that is  399 minutes and 100 minutes respectively.
Budget also has a wide range of variability and an amount of outliers . 
It is worth mentioning that votes have a large sandard deviation , 4370 to be exact and that might be because of an outlier that presents a huge difference with the rest. Possibly a small number of films had respectively huge number of votings compared to the rest of films. 





```{r}
table(film$genre)
```

```{r boxplot ,out.width='68%', fig.pos="H",fig.align='center', fig.cap="\\label{fig:boxplot}  IMDB rating by genre"}
ggplot(film, aes(genre,rating))+
  geom_boxplot(fill="SpringGreen3")+
  labs(x="Genre",y=" IMDB rating")+
  theme_light()
```


# Formal Data Analysis {#sec:FDA}


```{r} 
# Full model with all variables 

model_full <- glm ( Rate ~ year + genre + length + votes + budget , 
                    data= No_Rom_films , 
                    family = binomial(link = "logit"))
plot_model(model_full , show.values = TRUE , show.p = FALSE )

# now we'll check significance of those coeff 
# 95% CI 
confint( model_full) %>% 
   kable()%>% 
   kable_styling()


model_full %>% 
   coef() %>% 
   exp()
```

```{r}
# model using only significant explanatory

model2 <- glm ( Rate ~ genre + length + votes + budget , 
                    data= No_Rom_films , 
                    family = binomial(link = "logit"))

summary(model2)
```


# Conclusion and further task {#sec:Conclude}
